### 2010-02-21 ChrisnaSections.tex
Screw all of you.

\documentclass{article}
\usepackage{amsmath}
\begin{document}
\end{document}