### 2010-02-21 ChrisnaSections.tex

\documentclass{article}
\usepackage{amsmath}
\begin{document}
    \section{Standard Hierarchical Model}
        \subsection{Model Parameters}
        \subsection{Simulated Data}
    \section{Discussion}
    \begin{enumerate}
        \item Trying to get models to work on the actual data.
        \item What estimates we found for our data, what data didn't work, and
        why?
        \item Effect of plot size on parameters.
        \item Effect of covariates.
        \item Data with all plots versus data with every other plot.
        \item Best practices for further data gathering trips.
    \end{enumerate}
\end{document}