### 2010-02-23 kivaSections.tex
why am i writing ths?

%kkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkk
\documentclass[12pt]{article}
\usepackage{graphicx,amssymb}

\begin{document}
\section{Questions of investigation}
The North American river otter (\textit{Lontra canadensis}) is endemic to all of Minnesota; however, populations declined precipitously in the 19th and 20th centuries due to trapping and habitat degradation. In recent years their populations have significantly rebounded, especially in southeastern Minnesota where hunters would like to be able to trap otters again. In order for that to be possible governmental organizations must be able to monitor the population health. In addition, otter presence can indicate high quality aquatic habitat. Due to these reasons, there is interest in monitoring changes in otter abundance over time. Unfortunately, otters are not easy to observe. Thus, this study investigated the potential of using wintertime aerial surveys of otter tracks on rivers to make inferences about the abundance of otters and how it changes over time. We used the concepts from occupancy theory to develop models to estimate abundance.

\subsection{Data collection}
During the winters of 2003 and 2004 people flew over the Whitewater, Zumbrow, and Mississippi Rivers. When they saw what looked like an otter track, they pushed a button recording the GPS coordinates. They pushed the button every five seconds until the track ended. Flights occurred one, two, or three days after a snow event. In some cases, multiple observers flew over the same river on the same day. Surveys were conducted during multiple snow events throughout both winters. The rivers were then divided into 400 and 800 meter plots and the presence (1) or absence (0) of a waypoint in a plot on a given flight was determined. The Mississippi River was also divided into 1600 meter plots.

\subsection{Issues related to the data}
One major issue with respect to the data is that if a track occurred near the boundary between two plots, a person on one flight may have recorded the track in one site whereas someone on a later flight during that same snow event may have seen the same track but recorded it in the neighboring plot. So while the data indicate that the observers on the two flights disagreed in two plots, they saw the same track. Increasing the plot size should decrease the magnitude of this effect.

In addition, there is concern that the probability of an otter laying down a track in one site is not independent of the probability that an otter lays down a track in neighboring sites, violating a principle assumption of occupancy theory. That is, a single track can extend across multiple plots. This leads to dependence among the plots. It was important to quantify this dependence and find an optimum way to handle it.

\subsection{Descriptive statistics}

In 2003, three different observers flew over the rivers. Assuming that no tracks are laid down between individuals’ flights on the same day, all three people should have seen the same tracks. While there is some error in classifying which site the observer saw the actual track in, each observer should at least be seeing the same, or nearly the same, number of sites with tracks. However, we did not find that to be the case. If we look only at days that all three people observed, there appear to be major differences among the proportion of sites in which the scientists detected a track (Figure x). The difference in estimated na\"ive occupancy among the observers indicates that the scientists did not always correctly identify the tracks.

Correlation among the flight sequences is low. For example, average correlation between flights on the Mississippi in 2003 is 0.176 for 400 meter plots. Some of this is due to the difficulties in determining to which site each track belongs. However, we have also found quite a bit of variation in the data not necessarily due to these classification issues. This is difficult to quantify, as an algorithm for finding classification errors would be complicated. Correlation between flights is a bit higher for flights on the same day or flights during the same snow event, but still not as high as one might hope. The proportion of sites with tracks detected generally increases with days since snow, likely because the otters have a longer time to lay down a track as time after the snow event passes.

\subsection{False positive rates}
The lack of consistency among the flights is strong evidence that the data contain false positives where scientists think they see a track when it does not exist. Not accounting for such a process can lead to major biases; as the number of flights increases, the occupancy rate estimate will approach one (Royle and Link 2006). We used the observer as a covariate (by means of indicator variables) for both the probability of correctly identifying a track that exists ($p$) and the probability of identifying a track that does not exist ($E$). Including false positives introduces a duality whereby $(p = x, E = y, \psi = z)$ is as good a solution as $(p = y, E = x, \psi = 1-z)$. One possible way to address this problem is to assume $p\gg0.5$ and $E\ll0.5$.

Figure x caption:
The box plots show how the proportion of sites in which an observer saw tracks (“occupancy”) differs among the three observers. Only those dates in which all three observers flew are included.

\end{document}