### 2010-02-24 sarahDraft.tex
Added Chrisna's fixed equations

\documentclass{article}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage[hmargin=3cm, vmargin=3cm]{geometry}
\usepackage[all]{xy}
\begin{document}

\section{Overview of Occupancy Modeling}

	\subsection{What is Occupancy Rate?}

		Occupancy has two similar definitions.  It can be "the probability that a
randomly selected site or sampling unit in an area of interest is occupied by a 
species" [MacKenzie2006].  It also can be used to describe "the proportion of 
area, patches, or sample units that is occupied" [MacKenzie2006].  Occupancy is 
not the same as the abundance, or population size, of a species.  The number of
members of a species is not important, only whether or not at least one member of
the species is present in each site.

	While there is a difference between these two definitions of occupancy, in 
general the probability is unknown and the observed proportion is used as an 
estimate of the underlying probability.  Because of this, these two definitions
of occupancy are oftentimes used interchangably.

	The main reasons that people are interested in measuring occupancy are science
or conservation and management [MacKenzie2006].  This statistic can give insight 
into the way a system works.  It also can be used to monitor animals habitats 
and discover trends over time in animal populations.

	\subsection{Different Ways of Sampling Data}

		In order to determine occupancy rate, the data is often collected in the
 form of a presence-absence survey.  This means that observers visit sites and 
spend time within each one looking for the animals themselves, tracks, 
vocalizations, or some other indicator that the animal is present.  In some cases, 
camera traps and sound recording are used instead of human observers.

	The main problem with these methods of sampling data is that it is possible 
for an animal to occupy a site, but not be detected by the observer.  This means
that the site is recorded as absent, but there really is an animal there.  This 
is called a false negative.  Marking an animal present when there really is no
animal is also a possible problem, but is thought to be much less likely and is
often ignored.  We will discuss this more below.

	\subsection{Assumptions}

	There are some assumptions that are made in order to create models of 
occupancy.  Without these assumptions, the models become much more complicated.
One common assumption is no false positives, which means if a site is recorded as
present, this cannot be incorrect.  This is an assumption that we attempted to 
get rid of in our models.  A second common assumption is that presence and 
detection probabilties are constant across sites and surveys.  A third common 
assumption is that there is independence of detection between sites.  A fourth
common assumption is that the surveys are being conducted on a closed population.
This means that no animals leave the area or join the area during the time of the
surveys.

	Obviously, not all of these assumptions are always true.  In some cases, there
are established ways to deal with getting rid of an assumption in the models, 
and in others there are no common ways to get rid of the assumption.  Either way, 
it adds a level of complication to the model, and therefore we will only talk 
about dealing with the assumption of no false positives, and independence of 
detection across sites.

	\subsection{Types of Occupancy Models}

		All of the models that we looked at were single-species, single-season
models because this is what fit our data, but these models can be modified to
be used for multiple seasons or multiple species.

	The most simple way to measure occupancy is to simply divide the number 
of sites where an animal was found to be present by the total number of sites.  
This is called naive occupancy [MacKenzie2006].  It is generally accepted that 
this estimate is too low because animals are not detected at all sites where they
are present, which is the issue of false negatives.  Other occupancy models
attempt to do better than this simple estimate.

	One way to create more accurate models is to use a Hierarchical model as a 
way to think about the parameters that affect occupancy.  This allows us to take
into account that when an animal is not detected, it could either not be there, 
or it could be there and just was not detected.  This could either be due to
a false negative, where the observer did not see the indicator, or it could be 
due to the fact that the animal does occupy the site but there was no indication
of the animal at the site- for example, no track, no vocalization, etc.  \\
Here is a visualization of how the Hierarchical model works:
	\begin{displaymath}
		\xymatrix{          	   &                                  		 &                               		&\text{Site} \ar[dl] \ar[dr] & \\
				            	   &                                  		 & \text{Occupied} \ar[dl] \ar[d]		&                            & \text{Unoccupied} \ar[d] \\
				            	   & \text{Presence Indicated} \ar[dl] \ar[d]& \text{Presence Not Indicated} \ar[d]&                            & \text{Not Detected} \\
				   \text{Detected} & \text{Not Detected}              		 & \text{Not Detected}                  &                            &   }
	\end{displaymath}

From this diagram, you can see that we need separate parameters for the 
probability that a site is occupied, the probability that presence is indicated
given that a site is occupied, and the probability of detection given that 
presence is indicated and the site is occupied.  We call these \(\psi\),
\(\theta\), and \(p\), respectively.

\section{Actual data}
Stuff by Kiva.

	\subsection{Markov Chains and Testing the Independence Assumption}

		First we considered using Markov Chains to come up with a long term 
probability of detection, which we called \(r\).  We let \(p\) be the number of 
\(0\)'s followed by a \(0\) as a fraction of all successive pairs that start 
with a \(0\).  We let \(q\) be the number of \(1\)'s followed by a \(1\) as a 
fraction of all successive pairs that start with a \(1\).  These probabilities 
can be represented by the matrix \(M = \begin{bmatrix} p & 1-p \\ 1-q & q 
\end{bmatrix}\). The long term probability of detection can be found from this 
by finding \(\lim_{n \to \inf} M^n\), which returns \( \begin{bmatrix} 1-r & r 
\\ 1-r & r\end{bmatrix}\), where \(r\) is the probability of a \(1\), which is 
the long term probability of detection.

	We calculated \(p\), \(q\), and \(r\) for every flight in our data, using 
\(n = 500\).  We then realized that \(r\) does not answer the questions we are 
asking.  It does not take into account whether or not a site is occupied, or 
whether or not a track was laid down on that site, or any of the ideas behind    
the hierarchical model.  These are the parameters that we eventually want to 
find out something about, and the value of \(r\) cannot help us get there.  The 
ideas behind using Markov Chains did end up being useful, however.

	Although we rejected the idea of using \(r\) as a way to analyze the data, 
we realized that \(p\) and \(q\) could still be used to test the viability of 
the independence assumption.  If there is independece between sites, the 
difference of the proportions \(p\) and \(1-q\) should be approximately \(0\). 
This would mean that the probability that a \(0\) follows a \(0\) is about the 
same as the probability that a \(1\) follows a \(0\), meaning that it does not 
depend on what the observation is at the previous site.  Looking at boxplots of 
\(p - (1-q)\), we see that data from flights that record alternating sites may 
be more independent than data from flights that record all sites, which suggests
a possible benefit to using the alternating sampling method.  The boxplots for 
the Whitewater and Zumbro Rivers looked similar to the Mississippi:  \\
*INSERT BOXPLOT FROM STATUS REPORT HERE  \\
We found no obvious trend indicating one site size to be more independent than 
the others.

	While this is an interesting observation, we would like to be able to 
quantify it better by finding the distribution of our statistic.  This would 
allow us to make statements about how independent the data is, and if the 
difference in independence between different types of data is significant.   
This statistic seems hard to get, but we considered a modification.  Let \(S\) 
be the number of successive \(0\),\(1\)'s minus the number of successive \(1\),
\(1\)'s per flight.  Under the null hypothesis that each site is an independent 
Bernoulli trial with probability \(p\) we were able to compute the mean and 
standard deviation of \(S\), which we used to calculate a Z-score.
	In order to compute these values, we needed to find the expected value and
variance of each term in \(S\).  We let \(A = \text{\# of 0-1's}\), 
\(B = \text{\# of 1-1's}\) and \(p = P(``1")\text{ in the sequence}\).

	The mean ...
 	  % E[A]
    \begin{equation*}
        \begin{split}
        E[A]& =\sum_{i=1}^{n-1}P(X_i=0,X_{i+1}=1) \\
            & =\sum_{i=1}^{n-1}(1-p)p \\
            & =(n-1)(1-p)p \\
        \end{split}
    \end{equation*}

    % E[B]
    \begin{equation*}
        \begin{split}
        E[B]& =\sum_{i=1}^{n-1}P(X_i=1,X_{i+1}=1) \\
            & =\sum_{i=1}^{n-1}p*p \\
            & =(n-1)p^2
        \end{split}
    \end{equation*}

  % E[s-hat]
    \begin{equation*}
        \begin{split}
        E[\hat{s}]& =E[A]-E[B] \\
                  & =(n-1)\left((1-p)p-p^2\right)
        \end{split}
    \end{equation*}

	The standard deviation...
    % Var[A]
    \begin{equation*}
        \begin{split}
        Var[A]& =\sum_{i=1}^{n-1}Var[A]+2\sum_{i=1}^{n-2}Cov[I_i=(0,1),I_{i+1}=
                 (1,1)] \\
              & =(n-1)P((0,1))(1-P((0,1))) \\
              & \quad +2\sum_{i=1}^{n-2}\bigg(E[I_i=(0,1),I_{i+1}=(0,1)]-E[I_i=
                (0,1)]E[I_{i+1}=(0,1)]\bigg) \\
              & =(n-1)(1-p)p(1-(1-p)p)+2\sum_{i=1}^{n-2}\bigg(0-(1-p)p(1-p)p
                \bigg) \\
              & =(1-p)p(-1+n+5p-3np-5p^2+3np^2)
        \end{split}
    \end{equation*}

    % Var[B]
    \begin{equation*}
        \begin{split}
        Var[B]& =\sum_{i=1}^{n-1}Var[B]+2\sum_{i=1}^{n-2}Cov[I_i=(1,1),I_{i+1}=
                (1,1)] \\
              & =(n-1)P((1, 1))(1-P((1, 1))) \\
              & \quad +2\sum_{i=1}^{n-2}\bigg(E[I_i=(1,1),I_{i+1}=(1,1)]-E[I_i=
                (1,1)]E[I_{i+1}=(1,1)]\bigg) \\
              & =(n-1)p^2(1-p^2)+2(n-2)(p^3-p^4) \\
              & =(1-p)p^2(-1+n-5p+3np)
        \end{split}
    \end{equation*}

    % Cov[A, B]
    \begin{equation*}
        \begin{split}
        Cov[A,B]& =\sum_{i=1}^{n-1}\sum_{j=1}^{n-1}Cov[A,B] \\
                & =\sum_{i=1}^{n-2}Cov[A,B]+\sum_{i=1}^{n-2}Cov[B,A] \\
                & \quad +\sum_{i=1}^{n-1}Cov[X_i=0\text{ or }1,X_{i+1}=1] \\
                & =\sum_{i=1}^{n-2}E[AB]-E[A]E[B]+\sum_{i=1}^{n-2}E[BA]-
                  E[B]E[A] \\
                & \quad +\sum_{i=1}^{n-1}E[X_i=0\text{ or }1,X_{i+1}=1]-
                  E[B]E[A] \\
                & =(n-2)(p-1)^2p^2-(n-2)p^3(1-p)-(n-1)p^3(1-p)
        \end{split}
    \end{equation*}

    % Var[s-hat]
    \begin{equation*}
        \begin{split}
        Var[\hat{s}]& =Var[A-B] \\
                    & =Var[A]+Var[B]-2Cov[A,B] \\
                    & =(1-p)p(-1+n+8p-4np-20p^2+12np^2)
        \end{split}
    \end{equation*}

	Simulations of \(S\) show that the distribution of \(S\) is roughly normal.  
So using our calculations for the mean and the standard deviation, we computed
a Z-score for each flight, where we use a plug-in estimate for \(p\) to be the 
proportion of \(1\)'s per flight.  We said that Z-scores with an absolute value 
greater than \(2\) suggest that the assumption of independence between sites is 
violated for that flight.  We found that, in general, alternating sites are more
"independent" than all sites as measured by this statistic.  However, alternating
sites do not guarentee independence.  For instance, six of eleven flights along 
the Zumbro River during 2003 still have significant Z-scores with alternating
sites.  There were no noticable trends for this statistic with respect to 
observer, days since snow, snow event, or plot size.

*FIGURES OR TABLES??? WHICH ONES?

	We did find interesting trends with regard to sampling method.  We think 
that the lack of independence results from strings of ones where an otter left a
track across multiple sites.  When alternating sites are used, the strings are 
cut in half and the occurrence of \(1\),\(1\)'s is lower while the occurence of 
\(0\),\(1\)'s stays about the same.  It is noteworthy that virtually all the 
violations of independence are in the direction of more \(1\),\(1\)'s than would
be expected.  This reinforces our thoughts about successive \(1\)'s representing
the same otter.  We will discuss later whether this violation of independence 
actually matters in our model, and how this might affect the estimation of 
actual occupancy rates.

\section{Theoretical Stuff}

	\subsection{Bayesian Statistics}

	   	Bayes' Theorem: \(P(\theta|X) = P(X|\theta)P(\theta) / P(X)\) \\

	Usually, the mean of the posterior distribution is then used as a point 
estimate for the parameter.

Therefore, the posterior distribution \( \propto \) the likelihood times the prior.
    	Useful for hierarchical modeling: look at section 3.3 in book. \\
    	Can use Markov Chain Monte Carlo (MCMC) to calculate posterior 
			distribution- bimodal issues.

	\subsection{BRugs/R} - Chrisna

\end{document}