### 2010-02-24 sarahDraft.tex
Draft

\documentclass{article}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage[hmargin=3cm, vmargin=3cm]{geometry}
\usepackage[all]{xy}
\begin{document}

\section{Overview of Occupancy Modeling}

	\subsection{What is Occupancy Rate?}

		Occupancy has two different definitions.  It can mean "the probability that a
randomly selected site or sampling unit in an area of interest is occupied by a 
species" [MacKenzie2006].  It also can be used to describe "the proportion of 
area, patches, or sample units that is occupied" [MacKenzie2006].  Occupancy is 
not the same as the abundance, or population size, of a species.  The number of
members of a species is not important for either definition of occupancy.

	While there is a difference between these two definitions of occupancy, in 
general the probability is unknown and the observed proportion is used as an 
estimate of the underlying probability.  Because of this, these two definitions
of occupancy are oftentimes used interchangeably.

	The main reasons that people are interested in measuring occupancy are science
or conservation and management [MacKenzie2006].  This statistic can give insight 
into the way a system works.  It also can be used to monitor animals habitats 
and discover trends over time in groups of animals.

	\subsection{Different Ways of Sampling Data}

		In order to determine occupancy rate, the data is often collected in the
 form of a presence-absence survey.  This means that observers visit sites and 
look for the animals themselves, tracks, vocalizations, or some other indicator 
that the animal is present.  In some cases, camera traps and sound recording are 
used instead of human observers. [MacKenzie 2006]

	The main problem with these methods of sampling data is that it is possible 
for an animal to occupy a site, but not be detected by the observer.  This means
that the site is recorded as unoccupied, but there really is an animal there.  
This is called a false negative.  Marking an animal present when there really is
no animal is also a possible problem, but is thought to be much less likely and 
is often ignored.  We will discuss this issue more below.

	\subsection{Assumptions}

	There are some assumptions that are made in most models of occupancy.  [MacKenzie 2006].
Without these assumptions, the models become much more complicated.  One 
common assumption is no false positives, which means if a site is recorded as
occupied we say it must actually be occupied.  A second common assumption is that
presence and detection probabilities are constant across sites and surveys.  This 
simplifies the estimation of each parameter.  A third common assumption is that 
there is independence of detection between sites.  A fourth common assumption is 
that the surveys are being conducted on a closed population.  This means that no 
animals leave the area or join the area during the time of the surveys.

	Obviously, not all of these assumptions are always true, but they are often 
necessary in order to get a model that is simple enough to use.  There often are
possible ways to get rid of these assumptions, but it complicates the model.  In
this paper, we will only talk about attempting to get rid of the assumption of no
false positives, and the assumption of independence of parameters across sites.

	\subsection{Types of Occupancy Models}

		All of the models discussed in this paper are single-species, single-season
models.  It is possible to modify all of these models to be used for multiple-seasons
or multiple-species, but it would make them more complicated.

	The most simple way to measure occupancy is to simply divide the number 
of sites where an animal was found to be present by the total number of sites.  
This is called naive occupancy [MacKenzie2006].  It is generally accepted that 
this estimate is too low because animals are not detected at all sites where they
are present, which is the issue of false negatives mentioned above.  Other 
occupancy models attempt to do better than this simple estimate.

	One way to create more accurate models is to use a Hierarchical model as a 
way to think about the parameters that affect occupancy.  This allows us to take
into account that when an animal is not detected, it could either not be there, 
or it could be there and just was not detected.  This could either be due to
a false negative, where the observer did not see the indication- for example they 
did not see the animal, track, etc.- or it could be due to the fact that the 
animal does occupy the site but there was no indication of the animal at the 
site- for example, no track, no vocalization, etc.  \\
Here is a visualization of how the Hierarchical model works:
	\begin{displaymath}
		\xymatrix{          	   &                                  		 &                               		&\text{Site} \ar[dl] \ar[dr] & \\
				            	   &                                  		 & \text{Occupied} \ar[dl] \ar[d]		&                            & \text{Unoccupied} \ar[d] \\
				            	   & \text{Presence Indicated} \ar[dl] \ar[d]& \text{Presence Not Indicated} \ar[d]&                            & \text{Not Detected} \\
				   \text{Detected} & \text{Not Detected}              		 & \text{Not Detected}                  &                            &   }
	\end{displaymath}

From this diagram, it is apparent that we need separate parameters for the 
probability that a site is occupied, the probability that presence is indicated
given that a site is occupied, and the probability of detection given that 
presence is indicated and the site is occupied.  We will call these \(\psi\),
\(\theta\), and \(p\), respectively.

\section{Actual data}
Stuff by Kiva.

	\subsection{Markov Chains and Testing the Independence Assumption}

		First we considered using Markov Chains to come up with a long term 
probability of detection, which we called \(r\).  We let \(p\) be the number of 
\(0\)'s followed by a \(0\) as a fraction of all successive pairs that start 
with a \(0\).  We let \(q\) be the number of \(1\)'s followed by a \(1\) as a 
fraction of all successive pairs that start with a \(1\).  These probabilities 
can be represented by the matrix \(M = \begin{bmatrix} p & 1-p \\ 1-q & q 
\end{bmatrix}\). The long term probability of detection can be found from this 
by finding \(\lim_{n \to \inf} M^n\), which returns \( \begin{bmatrix} 1-r & r 
\\ 1-r & r\end{bmatrix}\), where \(r\) is the probability of a \(1\), which is 
the long term probability of detection.

	We calculated \(p\), \(q\), and \(r\) for every flight in our data, using 
\(n = 500\).  We then realized that \(r\) is actually not so useful in analyzing
our data.  It does not take into account whether or not a site is occupied, or 
whether or not a track was laid down on that site, or any of the ideas behind    
the hierarchical model.  These are the parameters that we eventually want to 
find out something about, and the value of \(r\) cannot help us get there.  The 
idea to use Markov Chains was still useful, however.

	Although we rejected the idea of using \(r\) as a way to analyze the data, 
we realized that \(p\) and \(q\) could be used to test the viability of 
the independence assumption.  If there is independence between sites, the 
difference of the proportions \(p\) and \(1-q\) should be approximately \(0\). 
This would mean that the probability that a \(0\) follows a \(0\) is about the 
same as the probability that a \(1\) follows a \(0\), meaning that it does not 
depend on what the observation is at the previous site.  Looking at boxplots of 
\(p - (1-q)\), we see that data from flights that record alternating sites may 
be more independent than data from flights that record all sites, which suggests
a possible benefit to using the alternating sampling method.  The boxplots for 
the Whitewater and Zumbro Rivers looked similar to the Mississippi:  \\
*INSERT BOXPLOT FROM STATUS REPORT HERE OF p-(1-q) Mississippi \\
We found no obvious trend indicating one site size to be more independent than 
the others.

	While this is an interesting observation, we would like to be able to 
quantify it better by finding the distribution of our statistic.  This would 
allow us to make statements about how independent the data is, and if the 
difference in independence between different types of data is significant.   
This statistic seems hard to get, but we considered a modification.  Let \(S\) 
be the number of successive \(0\),\(1\)'s minus the number of successive \(1\),
\(1\)'s per flight.  Under the null hypothesis that each site is an independent 
Bernoulli trial with probability \(p\) we were able to compute the mean and 
standard deviation of \(S\), which we used to calculate a Z-score.
	In order to compute these values, we needed to find the expected value and
variance of each term in \(S\).  We let \(A = \text{\# of 0-1's}\), 
\(B = \text{\# of 1-1's}\) and \(p = P(``1")\text{ in the sequence}\).

In order to find \(E[\hat{s}]\), we first needed to find \(E[A]\) and \(E[B]\).
 	  % E[A]
    \begin{equation*}
        \begin{split}
        E[A]& =\sum_{i=1}^{n-1}P(X_i=0,X_{i+1}=1) \\
            & =\sum_{i=1}^{n-1}(1-p)p \\
            & =(n-1)(1-p)p \\
        \end{split}
    \end{equation*}

    % E[B]
    \begin{equation*}
        \begin{split}
        E[B]& =\sum_{i=1}^{n-1}P(X_i=1,X_{i+1}=1) \\
            & =\sum_{i=1}^{n-1}p*p \\
            & =(n-1)p^2
        \end{split}
    \end{equation*}

  % E[s-hat]
    \begin{equation*}
        \begin{split}
        E[\hat{s}]& =E[A]-E[B] \\
                  & =(n-1)\left((1-p)p-p^2\right)
        \end{split}
    \end{equation*}

The standard deviation was a little more difficult.  
    % Var[A]
    \begin{equation*}
        \begin{split}
        Var[A]& =\sum_{i=1}^{n-1}Var[A]+2\sum_{i=1}^{n-2}Cov[I_i=(0,1),I_{i+1}=
                 (1,1)] \\
              & =(n-1)P((0,1))(1-P((0,1))) \\
              & \quad +2\sum_{i=1}^{n-2}\bigg(E[I_i=(0,1),I_{i+1}=(0,1)]-E[I_i=
                (0,1)]E[I_{i+1}=(0,1)]\bigg) \\
              & =(n-1)(1-p)p(1-(1-p)p)+2\sum_{i=1}^{n-2}\bigg(0-(1-p)p(1-p)p
                \bigg) \\
              & =(1-p)p(-1+n+5p-3np-5p^2+3np^2)
        \end{split}
    \end{equation*}

    % Var[B]
    \begin{equation*}
        \begin{split}
        Var[B]& =\sum_{i=1}^{n-1}Var[B]+2\sum_{i=1}^{n-2}Cov[I_i=(1,1),I_{i+1}=
                (1,1)] \\
              & =(n-1)P((1, 1))(1-P((1, 1))) \\
              & \quad +2\sum_{i=1}^{n-2}\bigg(E[I_i=(1,1),I_{i+1}=(1,1)]-E[I_i=
                (1,1)]E[I_{i+1}=(1,1)]\bigg) \\
              & =(n-1)p^2(1-p^2)+2(n-2)(p^3-p^4) \\
              & =(1-p)p^2(-1+n-5p+3np)
        \end{split}
    \end{equation*}

    % Cov[A, B]
    \begin{equation*}
        \begin{split}
        Cov[A,B]& =\sum_{i=1}^{n-1}\sum_{j=1}^{n-1}Cov[A,B] \\
                & =\sum_{i=1}^{n-2}Cov[A,B]+\sum_{i=1}^{n-2}Cov[B,A] \\
                & \quad +\sum_{i=1}^{n-1}Cov[X_i=0\text{ or }1,X_{i+1}=1] \\
                & =\sum_{i=1}^{n-2}E[AB]-E[A]E[B]+\sum_{i=1}^{n-2}E[BA]-
                  E[B]E[A] \\
                & \quad +\sum_{i=1}^{n-1}E[X_i=0\text{ or }1,X_{i+1}=1]-
                  E[B]E[A] \\
                & =(n-2)(p-1)^2p^2-(n-2)p^3(1-p)-(n-1)p^3(1-p)
        \end{split}
    \end{equation*}

    % Var[s-hat]
    \begin{equation*}
        \begin{split}
        Var[\hat{s}]& =Var[A-B] \\
                    & =Var[A]+Var[B]-2Cov[A,B] \\
                    & =(1-p)p(-1+n+8p-4np-20p^2+12np^2)
        \end{split}
    \end{equation*}

	Simulations of \(S\) show that the distribution of \(S\) is roughly normal.  
So using our calculations for the mean and the standard deviation, we computed
a Z-score for each flight, where we use a plug-in estimate for \(p\) to be the 
proportion of \(1\)'s per flight.  We said that Z-scores with an absolute value 
greater than \(2\) suggest that the assumption of independence between sites is 
violated for that flight.  We found that, in general, alternating sites are more
"independent" than all sites as measured by this statistic.  However, alternating
sites do not guarantee independence.  For instance, six of eleven flights along 
the Zumbro River during 2003 still have significant Z-scores with alternating
sites.  There were no noticeable trends for this statistic with respect to 
observer, days since snow, snow event, or plot size.

*FIGURES OR TABLES??? WHICH ONES? from status report compilation

	We think that the lack of independence results from strings of ones where an 
otter left a track across multiple sites.  When alternating sites are used, the 
strings are cut in half and the occurrence of \(1\),\(1\)'s is lower while the 
occurrence of \(0\),\(1\)'s stays about the same.  It is noteworthy that 
virtually all the violations of independence are in the direction of more 
\(1\),\(1\)'s than would be expected.  This reinforces our thoughts about 
successive \(1\)'s representing the same otter.  We will discuss later whether 
this violation of independence actually matters in our model, and how this might
affect the estimation of actual occupancy rates.

\section{Theoretical Stuff}

	\subsection{Bayesian Statistics}

	   	Bayes' Theorem: \(P(\theta|X) = P(X|\theta)P(\theta) / P(X)\) \\

	This states that the posterior distribution \( \propto \) the likelihood 
\(\times\) the prior.
	Usually, the mean of the posterior distribution is then used as a point 
estimate for the parameter.

	We can use Bayesian statistics in the Hierarchical model as long as we set a
prior for every parameter that we include.  We can then use Markov Chain Monte 
Carlo (MCMC) to find a posterior distributions for every parameter.  We run into 
a problem of duality with the probability of correct detection \(p\), and the 
probability of false detection \(E\).  These two variables have bimodal posterior
distributions, because \(p = x, E = y, \psi = z\) is the same as \(p = y, E = x, 
\psi = 1 - z\).  Since we know that is reasonable for \(p\) to be greater than \(E\),
we solved this problem by restricting the distribution of \(p\) to the range
\(0.6 \text{ to } 1\), and \(E\) to \(0 \text{ to } 0.4\).

\end{document}