### 2010-06-08 NewPaper.tex
First draft.

\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{hyperref}

\renewcommand{\bibsection}{\section{References}}

\begin{document}
\addtolength{\textheight}{+0.18in}

\newcommand{\lf}{\left\lfloor}
\newcommand{\rf}{\right\rfloor}
\newcommand{\dis}{\displaystyle}
\newcommand{\sums}[2]{\sum_{\stackrel{{\scriptstyle {#1}}}{{#2}}}}
\newcommand{\prods}[2]{\prod_{\stackrel{{\scriptstyle {#1}}}{{#2}}}}

\long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

\def\prbox{\hfill\rule{1.2ex}{1.2ex}\vspace{.2in}}

\newenvironment{proof}{\noindent{\bf Proof }}{\prbox}

\def\Reals{\hbox{\rm I\kern-.18em R}}
\def\reals{\Reals}
\def\Complexes{\hbox{\rm C\kern-.43em
       \vrule depth 0ex height 1.4ex width .05em\kern.41em}}
\def\complexes{\Complexes}

\def\field{\hbox{\rm I\kern-.18em F}} %symbol for field
\def\Naturals{\hbox{\rm I\kern-.17em N}}
\def\naturals{\Naturals}
\def\integers{\hbox{\rm Z\kern-.3em Z}}
\def\hh{\hrule height0.9pt width1.1em}
\def\vv{\vrule width0.8pt depth0.12em height0.92em}
\def\square{\vbox{\kern0.15em\hh\kern0.9em\hh\kern-1.05em
            \hbox{\vv\kern0.93em\vv}}}
\def\Lv{L_{\alpha} }
\def\Lvp{L_{\alpha'} }
\def\sm{{\mbox{\boldmath $\sigma$}_m}}
\def\MMv{R_{\mbox{\boldmath $\sigma$}_v}}
\def\tMMv{\widetilde{R}_{\mbox{\boldmath $\sigma$}_v}}
\def\MMm{R_{\mbox{\boldmath $\sigma$}_m}}
\def\tMMm{\widetilde{R}_{\mbox{\boldmath $\sigma$}_m}}
\def\MMj{R_{\mbox{\boldmath $\sigma$}_j}}
\def\sn1{{\mbox{\boldmath $\sigma$}_{n-1}}}
\def\snv{{\mbox{\boldmath $\sigma$}_v}}

\hyphenation{eigen-value so-lu-tions heur-is-tics mod-eled de-note
prob-a-bil-ity start-ing}

\LARGE\bf
\noindent
\textsf{A Bayesian hierarchical occupancy model for track surveys conducted in
a series of linear, spatially correlated, sites}

\rm\large
\vspace{2ex}
\noindent
\textsf{[short title: Bayesian occupancy model for track surveys]}

\rm\large
\normalsize

\vspace{4ex}
\small\bf
\noindent
\textsf{Chrisna Aing\footnotemark[1]\footnotetext[1]{Department of Mathematics,
Carleton College, Northfield, MN 55057, USA}, Sarah Halls\footnotemark[1], Kiva
Oken\footnotemark[1], Robert Dobrow\footnotemark[1]\symbolfootnote[1]
{Correspondence author: E-mail: rdobrow@carleton.edu},
John Fieberg \footnotemark[2]\footnotetext[2]{Minnesota Department of Natural
Resources}}

\normalsize\rm
\vspace{2ex}
\noindent
\rule{5in}{.4mm}

\vspace{3ex}
\begin{quote}
\noindent
{\bf Summary}
\begin{enumerate}
    \item In creating an occupancy model for river otters ({\em Lontra
        canadensis}) in southern Minnesota we develop an analytical approach
        that can be used in conjunction with survey data collected in a set of
        linear aerial survey sites.
    \item Our approach incorporates several recent developments in occupancy
        modeling, including the prevalence of false positives, spatial
        correlation, and repeated sampling across distinct observers and across
        daily snow events.
    \item We test our method using simulated and real data. Simulations across
        a wide range of parameters show the model to be robust, both with
        regard to false positive and negatives, and with regard to the type of
        spatial correlation observable in our data.  Confidence intervals for
        both the false positive and false negative rates are remarkably small
        given the relative paucity of the data.
    \item Our approach lends itself to investigating the optimal allocation of
        sampling effort and we make several recommendations for future studies.
\end{enumerate}

\vspace{2ex}
\noindent
{\bf Key-words:} Hierarchical modeling, occupancy, otter, simulation, spatial
correlation, track survey
\end{quote}

\normalsize
\section{Introduction}

\textbf{[Should we add all of these sources to BibTeX?]}

Wildlife populations are notoriously difficult to monitor, owing to challenges
of detecting individuals and the costs of surveying difficult to reach habitat.
A variety of approaches have been developed for estimating detection
probabilities in wildlife surveys, include double sampling, removal methods,
multiple observers, mark-recapture, distance sampling, and regression models
\citep{Lancia2005}.

\textbf{[The first citation below needs Thompson et al. 2004.]}

These methods, however, typically require identifying and possibly marking
unique individuals, and are therefore difficult to reliably implement in large
scale monitoring efforts or with elusive species \citep{Stanley2005,
Johnson2008}.  As a result, survey designs that record presence or absence of
animal sign (e.g., tracks, scent marks, feces) remain popular among natural
resource agencies \citep{Stanley2005}.  A flurry of recent research has been
devoted to developing methods, most often based on a survey design in which
sites are repeatedly visited, to correct for uncertain detection in
presence-absence surveys; the collection of methods for monitoring populations
using repeated presence-absence surveys has come to be known as ``occupancy
modeling'' \citep{MacKenzie2006, Royle2008}.

The primary purpose of this paper is to illustrate the development of an
occupancy modeling approach to analyzing river otter (\textit{Lontra
canadensis}) track survey data collected in the winter following snowfall
events.  Our approach will incorporate several recent developments in occupancy
modeling, including the prevalence of false detections \citep{Royle2006},
spatial correlation \citep{Magoun2007}, and a model for how detection
probabilities depend on time since a snowfall event \citep{Stanley2005}.

The data we consider were collected by the Minnesota Department of Natural
Resources (MNDNR) in 2003 using helicopter surveys with multiple observers,
multiple sampling intervals (representing unique snowfall events), and multiple
days within each sampling interval (days since the last snowfall event).
Flight paths followed the main channels of river systems, which were later
divided into a series of spatially contiguous quadrats (hereafter, we will
refer to these quadrats as ``sites'').  Due to potential for clustering of
individual home ranges in space and movement across site boundaries, we
expected the response, presence of a track (Y/N), to be spatially correlated
among sites.  In addition, sites were more likely to contain tracks if they
were sampled several days after a snowfall event.  Lastly, observer-to-observer
variability was significant, likely owing to measurement error associated with
the post-hoc creation of site boundaries and the difficulty of correctly
identifying otter tracks \citep{Evans2009}.  Thus, these data serve as a
useful case study to illustrate methods for handling several complications that
may be present in repeated presence-absence survey applications.

In addition to fitting a series of models to the Minnesota otter data, we use
extensive simulations to evaluate the robustness and precision of model-based
estimates of occupancy probabilities under this sampling design.  To promote
learning, and to facilitate adaption to other applications, we implement our
approach using open source software, Program R \citep{R2009} and WinBUGS
\citep{Lunn2000}, with the R package BRugs \citep{Thomas2006} to communicate
between the two software platforms.  Lastly, we show how simulations can be
used to compare different sampling designs, and we make recommendations
regarding allocation of sampling effort in future surveys.

\section{Materials and Methods}

    \subsection{Otter survey data collection}

    \textbf{[Need to cite Danny's dissertation (see next line).]}

    Details of the survey are given in \citet{Martin2003} and Martin (cite
    Danny's dissertation), but generally, aerial snow-track surveys were
    conducted using a Bell OH-58A+ helicopter following snowfalls of greater
    than 2.5 cm in depth.  The data discussed in this paper were collected in
    2003 from the Mississippi River.  There were five unique snowfall events.
    Flights occurred one, two or three days after each unique snowfall event,
    although they did not occur on every day.  There were three observers, and
    in some cases multiple observers flew over the river in different flights
    on the same day.  Observers collected a Global Positioning System waypoint
    upon encountering a track, and continued to log waypoints every five
    seconds thereafter if a track remained present.  After these waypoints were
    logged, three different datasets were created.  The section of the
    Mississippi River that was flown was divided into 402 meter sites, 804
    meter sites, and 1608 meter sites.  The presence (1) or absence (0) of a
    waypoint in each site on a given flight was determined and recorded.

    \subsection{Model development and likelihood formulation}

    Fundamentally, these data consist of a set of correlated binary (0, 1)
    responses arising from a set of biological and observational processes.  To
    construct the likelihood, we need to think about the different ways of
    arriving at a 1 (recorded presence of a track) or a 0 (absence of an
    apparent track).

    Let:

    $\psi_s=P(\mbox{site } s \mbox{ is occupied})$

    $\theta_s=P(\mbox{a track is laid down in site } s \mbox{ during the course
    of a day} | \mbox{site } s \mbox{ is occupied})$

    $p_k=P(\mbox{Observer } k \mbox{ detects a track} | \mbox{a track is
    present}).$

    $e_k=P(\mbox{Observer } k \mbox{ falsely records a track}| \mbox{there is
    not a track})$

    \textbf{[This is ugly :(.]}

    We assume the $\theta$'s are independent across days and constant across
    sampling instances, and that $p_k$ and $e_k$ are constant across sites and
    sampling instances. We define a site as being occupied if an otter uses the
    site for habitat during the winter.

    These assumptions together imply that the probability site $i$ contains a
    track, and this track is correctly detected by observer $l$ $k$ days after
    a snowfall event is given by: \[\psi_s [1-(1- \theta_s)^j] p_k\text{.}\]
    Here, $1-(1- \theta_s)^j$ is the probability that at least one track is
    laid down between the snowfall event and the survey date. This implies that
    the probability of a track being laid down on a given day increases with
    time since snowfall events.  There are five other possible outcomes, three
    of which lead to a 0 (recorded absence) and two more of which lead to a 1
    (recorded presence); all 6 outcomes are depicted in Figure 1. [Note:  I
    will add a figure with the hierarchical process and the probabilities of
    each event at the bottom of the figure].

    The four parameters described above: (i) occupancy rate $\psi$, (ii) a
    track-laying parameter $\theta$, (iii) detection probability $p$, and the
    false positive rate $e$, are the central parameters of our model.

    \textbf{[Should we include the paragraph below?]}

    Our data indicate a strong potential for false positives, consistent with
    other otter studies \citep{Evans2009}. The false positive parameter $e$ is
    the probability that a track is ``observed'' given that no track exists in
    the site on the given day. Thus false positives can occur in unoccupied
    sites or in an occupied site without a track. Both the correct detection
    ($p$) and false detection ($e$) probabilities depend on the observer, so
    that $p=p_l$ and $e=e_l$ for some observer \(l\).

    We omit the details but the likelihood function constructed from the above
    processes is described as follows. Let $S$  be the number of snow events in
    the season. For each snow event $i$, let $N_i$ be the number of sites
    sampled and $D_i$ be the number of days that were sampled from that snow
    event, which is typically no greater than three. And for each snow event
    $i$ and sampled snow day $j$, let $O_{ij}$ be the number of observers who
    surveyed the river, which is also typically no greater than three.  We
    denote the data by the vector ${\bf X}$. If we assume that occupancy and
    track-laying probabilities are independent and constant across sites, then
    the model likelihood is:

    \textbf{[No more than 2 days were sampled for any snowfall event
    on MS in 2003, but the flights ranged from 1 to 3 days after the actual
    event, so sometimes only days 2 and 3. I think for both days and observers
    it might be a safer bet to use sets instead of these integers.]}

    \begin{eqnarray*}
        L(\psi, \theta, \mathbf{p}, \mathbf{e}|\mathbf{X})& =&
        \prod_{i=1}^S \prod_{j=1}^{D_i} \prod_{k=1}^{O_{ij}} \left[ \left( \psi
        [(1-(1-\theta)^j) p_k +(1-\theta)^j) e_k ] + (1-\psi)e_k
        \right)^{n_{ijk}} \right. \\
        && \times \left( \psi[(1-(1-\theta)^j)(1-p_k)+(1-\theta)^j(1-e_k)]
        \right. \\
        && \left. \left. +(1-\psi)(1-e_k) \right)^{N_i-n_{ijk}}\right],
    \end{eqnarray*}
    where $n_{ijk}$ is the number of sites that record a 1 for the survey that
    was conducted $j$ days after snow event $i$ by observer $k$.

    \textbf{[The first line of the equation above crosses the margin.]}

    The construction of the above likelihood, under a maximum-likelihood
    framework, assumes a small sample of sites from a theoretically infinite
    sample space, independence across sites and independence between observers.
    While the latter is reasonable, the first two assumptions are almost surely
    not. In order to make best use of flight time, the entire section of the
    river was sampled. In addition, otters tend to travel for long distances
    [?] in the snow and a track at site $i$ is positively correlated with the
    existence of a track at site $i+1$. We quantified this association with
    standard runs tests for independence in a sequence of Bernoulli trials and
    found that for the 402-meter designs the null hypothesis of independence
    was strongly rejected in most surveys.  Correlations decreased, as would be
    expected, with larger plot sizes, but the positive correlation was often
    still statistically significant.  We address the spatial correlation issue
    below.

    \textbf{[The paragraph below is also mentioned in the introduction.]}

    The hierarchical model is more easily implemented via Markov Chain Monte
    Carlo (MCMC) using R, WinBUGS, and BRugs (see Appendix). The MCMC approach
    generates a Markov Chain to approximate the otherwise intractable posterior
    distribution.  The base model converges within about ten minutes of
    computing time on a standard desktop PC and 5,000 iterations of the MCMC
    chain.  WinBUGS is popular free and open-source software for MCMC. BRugs is
    an open-source package for R that contains the OpenBUGS library.  This
    allows users most of WinBUGS' functionality from within R.  For background
    on Bayesian hierarchical modeling and MCMC see \citet{Royle2008} and
    \citet{Ntzoufras2009}.

    We first constructed a simpler model that assumes independence across sites
    (but not among observers. We used a Uniform$(0,1)$ prior distribution for
    \(\psi\) and \(\theta\).  The \(p_i\)'s are from a beta\((a, b)\)
    distribution and the \(e_i\)'s are from beta\((c, b)\) distribution, where
    \(a\), \(b\) and \(c\) each come from gamma\((1, 0.1)\) distributions.
    After running the iterations of the MCMC chain, the model returns
    approximations of the posterior distribution for each variable.

    \citet{Royle2006} point out a problem with the ``dual role'' of $p$ and
    $e$ that occurs when false positives are added to a likelihood-based model.
    This results from the lack of contextual information.  In particular, for
    our model if $\theta =1$ then the likelihood above has the symmetrical
    property that $L(\psi, \theta, p, e) = L(1-\psi, \theta, e, p)$. This gives
    rise to bimodal posterior distributions. The suggested solution is to
    restrict the parameter space so that $p >> e$.

    In our Bayesian implementation we restricted the support for $p$ and $e$ to
    $(.6, 1)$ and $(0, .3)$, respectively, and this worked nicely. Note that
    for general $\theta$ the duality in the likelihood is no longer as simple
    as described above. But the problem with unspecified parameters and
    bimodality persists.

    Finally, we needed to consider the spatial dependence the data displayed.
    To do this, we considered conditional autoregressive (CAR) models that
    explicitly take into account such correlation. We considered one such model
    with a CAR prior placed only on occupancy, $\psi$, and another model with
    CAR priors placed on both $\psi$ and the track-laying process, $\theta$.
    The conditional variance of the CAR distributions, $\tau^2$, had a
    gamma$(0.5,0.0005)$ prior.

    \subsection{Simulation work}

    To test the ability of the model to reliably and accurately converge,
    simulated data sets with and without spatial correlation were generated and
    analyzed by the model. Three types of data were generated: data
    uncorrelated in \(\psi\) and \(\theta\), data with spatial correlation in
    \(\psi\) and uncorrelated in \(\theta\), and data with spatial correlation
    in both \(\psi\) and \(\theta\).

    Simulated data without spatial correlation is generated with given values
    of $\psi$, $\theta$, and $p$ and $e$ for each observer.

    To determine the efficacy of the model over the range of possible \(\psi\)
    and \(\theta\) values, simulated datasets with different \(\psi\) and
    \(\theta\) values were run through the model. All other variables: numbers
    of plots (100), observers (3), snow events (5), and observations after each
    snow event (3), were fixed. The model's estimates were then compared to the
    parameters used to generate the simulated data.

    As discussed, it is reasonable to believe that if an otter occupies one
    site, the neighboring sites have a higher probability of occupancy because
    that same otter may occupy consecutive sites.  Such a correlation in the
    occupancy process is incorporated into the simulated data through an ARIMA
    process.  First, a latent autoregressive process, \(\epsilon_i,\)
    \(i=1,...,n,\) is generated with a given unconditional variance of
    \(\sigma^2\) and a given autocorrelation parameter \(a\), using the
    arima.sim function in R.  Then occupancy probabilities for each site are
    generated using [notation? bo?] \(\psi_i=\mbox{logit}(\epsilon_i+bo)\),
    \(bo\) is a parameter representing the mean of the logit-transformed
    \(\psi\)'s.  This ensures that the \(\psi\)'s are between \(0\) and \(1\)
    and further allows every site \(i\) to have a different \(\psi_i\) that is
    correlated with \(\psi_{i-1}\).  We can then approximate the overall
    occupancy rate using integration to calculate \(E[\psi]\).

    Simulated binary track presence data with correlation only in \(\psi\) were
    generated using the above ARIMA machinery. The various models' estimates
    were then compared to the paramaters used to generate the simulated data,
    and confidence intervals were examined.

    Because a single track can extend over multiple sites, we also simulated
    data that displayed correlation in both the track-laying and occupancy
    processes.  Correlation in the track-laying process is incorporated into
    the simulated data through the addition of the variable \(\alpha\) and the
    redefinition of \(\theta\).  In this simulation method, \(\theta\)
    represents the probability that an otter lays a track in plot \(i\), given
    that plot \(i\) is occupied and plot \(i-1\) does \textit{not} contain a
    track. If plot \(i\) is occupied and plot \(i-1\) \textit{does} contain a
    track, then the probability of an otter making a track in plot \(i\) is
    \(\theta+\alpha\). Note that the correlation in occupancy is still
    incorporated through the ARIMA process.

    \subsection{Analysis of data}

    \textbf{[Should we bother with this paragraph? It's not really correct
    anymore anyway.  Chrisna: For the actual data, it is correct.  I'm not sure
    we ever ran anything new on the actual data, and I'm not even sure if it's
    possible at the moment because I may not have finished that part of the
    programming.]}

    The model analyzed data for each plot size of each river. For all data, the
    model was run with a burn-in period of ten thousand iterations.  Then
    parameters were tracked for ten thousand additional iterations.  If the
    model converged on the first run of the data, no further analysis was done.
    If the model did not converge on the first run of the data, it was run a
    second time with a burn-in period of twenty thousand iterations and a
    tracking period of twenty thousand iterations.  No further analysis was
    done.  The parameters estimated were \(E\) and \(p\) for each observer,
    \(\psi\), and \(\theta\).

    \subsection{Simulation work to look at study design}

    \textbf{[Now aren't we more interested in confidence intervals and less
    interested in convergence? We will have to redo all of this work.  Chrisna:
    Yes, we are.  I think this is pretty obvious though -- the more data points
    you have, the smaller the confidence intervals will be.]}

    We used simulations of data to determine the most efficient sampling
    practices (fewest number of flights) that produce data that converge with
    accurate estimates in the model.  The model is run with a burn-in period of
    10,000 iterations followed by a tracking period of 10,000 iterations.  The
    values used for the parameters in the simulated data are similar to those
    found in the results of the actual data obtained by running the model.  All
    of the data that was simulated had spatial correlation in both \(\psi\) and
    \(\theta\).  In each simulation there were one hundred plots, \(\psi\) was
    .3 with an \(\alpha\) of .5, \(\theta\) was .4 with a \(\beta\) of .4, all
    \(p\)'s were .7, and all \(E\)'s were .05.

    The number of snow events, days flown after a snow event, and number of
    observers are varied in order to determine the best combination of these
    parameters that will result in the fewest flights and a successful run of
    the model.  Data collected using one, two, and three observers were
    simulated.  For each of these numbers of observers, data with one, two, and
    three snow events were simulated.  For each of these numbers of observers
    combined with each of these numbers of snow events, data with one, two, and
    three consecutive days since snow were simulated.  Most of these
    combinations returned accurate estimates and good convergence when run in
    the model.  For the combinations that did not, more data was simulated with
    the number of snow events increased to five and ten.  Simulating data with
    consecutive days since snow increased to five was also tried.

\section{Results}

    \subsection{Model efficacy}

    The efficacy of three models was tested with three types of simulated data.
    The three models are the basic hierarchical model, the basic model with a
    CAR prior distribution on \(\psi\), and the basic model with CAR prior
    distributions on both \(\psi\) and \(\theta\).  The three types of
    simulated data are data with fixed parameters \(\psi\) and \(\theta\), data
    with \(\psi\) generated by the previously described ARIMA process and
    \(\theta\) fixed, and data with both \(\psi\) and \(\theta\) generated by
    the ARIMA process.  The other parameters are held constant.  The number of
    sites is 100, the number of observers is three, the number of observations
    after each snow event is three, \(p = (0.7, 0.85, 0.9)\) and \(E = (0.05,
    0.10, 0.20)\).

    We state that the model converges well if the MCMC diagnostics show
    convergence of two simulated chains, the final posterior distributions are
    unimodal, and the simulation parameter values fall within two SE's of the
    model's estimates.  For the most part, \(SE[E] \approx 0.01\) and \(SE(p)
    \approx 0.02\).

    \textbf{[I have no idea how to put the simulation results into words.]}

    Regardless of the type of data given to each of the three models, they
    perform similarly.  The biggest variation between the three models concerns
    the amount of iterations required for the model to converge.  When given
    uncorrelated data, all three models take approximately the same amount of
    iterations to converge.  The number of iterations depends upon the values
    of \(\psi\) and \(\theta\).  As \(\psi\) approaches \(\theta\), the amount
    of time required for the model to converge tends to increase.  When given
    the other two types of data, with correlation in \(\psi\) and correlation
    in \(psi\) and \(\theta\) the two models with CAR prior distributions
    occasionally require additional burn-in to converge.

    \subsection{Analysis of data}

    % Mississippi River!
    \begin{table}
        \caption{Mississippi River. The data converged well at the 402m and
        804m plot sizes. At the 1608m plot size, \(E\), \(\psi\), and
        \(\theta\) did not converge well.}
        \label{Mississippi}
        \begin{center}
            \begin{tabular}{|l|l|l|l|}
                \hline
                \multicolumn{4}{|l|}{\textbf{Mississippi River}} \\
                \hline
                    & 402m & 804m & 1608m \\
                \hline
                \multirow{2}{*}{E[Danny]}
                    & 0.01954 & 0.04808 & 0.06885 \\
                    & (0.006515, 0.03663) & (0.01874, 0.08187) & (0.01778,
                    0.1383) \\
                \hline
                \multirow{2}{*}{E[John]}
                    & 0.05216 & 0.07771 & 0.13650 \\
                    & (0.027710, 0.07745) & (0.03881, 0.12100) & (0.03449,
                    0.2330) \\
                \hline
                \multirow{2}{*}{E[Tom]}
                    & 0.02668 & 0.06140 & 0.10230 \\
                    & (0.011990, 0.04622) & (0.02789, 0.10390) & (0.04287,
                    0.1796) \\
                \hline
                \multirow{2}{*}{p[Danny]}
                    & 0.65240 & 0.67340 & 0.74960 \\
                    & (0.548300, 0.75990) & (0.56760, 0.78290) & (0.60510,
                    0.8910) \\
                \hline
                \multirow{2}{*}{p[John]}
                    & 0.67770 & 0.77580 & 0.82540 \\
                    & (0.583500, 0.77280) & (0.66060, 0.88690) & (0.71390,
                    0.9326) \\
                \hline
                \multirow{2}{*}{p[Tom]}
                    & 0.53670 & 0.54220 & 0.65330 \\
                    & (0.501200, 0.61300) & (0.50160, 0.62850) & (0.51530,
                    0.8104) \\
                \hline
                \multirow{2}{*}{\(\psi\)}
                    & 0.59380 & 0.69370 & 0.75820 \\
                    & (0.439000, 0.75990) & (0.51800, 0.87380) & (0.54110,
                    0.9300) \\
                \hline
                \multirow{2}{*}{\(\theta\)}
                    & 0.15280 & 0.20730 & 0.27690 \\
                    & (0.111900, 0.19780) & (0.14580, 0.27390) & (0.19030,
                    0.3716) \\
                \hline
            \end{tabular}
        \end{center}
    \end{table}

    % Whitewater River!
    \begin{table}
        \caption{Whitewater River. While the 402m data converged well, \(E\),
        \(\psi\), and \(\theta\) did not converge well for the 804m data.}
        \label{Whitewater}
        \begin{center}
            \begin{tabular}{|l|l|l|}
                \hline
                \multicolumn{3}{|l|}{\textbf{Whitewater River, 2003}} \\
                \hline
                    & 402m & 804m \\
                \hline
                \multirow{2}{*}{E[Danny]}
                    & 0.05608 & 0.0555 \\
                    & (0.02474, 0.09554) & (0.002859, 0.1390) \\
                \hline
                \multirow{2}{*}{E[John]}
                    & 0.16570 & 0.1931 \\
                    & (0.10130, 0.23510) & (0.065910, 0.3384) \\
                \hline
                \multirow{2}{*}{E[Tom]}
                    & 0.15350 & 0.2241 \\
                    & (0.09758, 0.21660) & (0.110000, 0.3623) \\
                \hline
                \multirow{2}{*}{p[Danny]}
                    & 0.66990 & 0.6985 \\
                    & (0.56080, 0.79040) & (0.573100, 0.8366) \\
                \hline
                \multirow{2}{*}{p[John]}
                    & 0.92040 & 0.9877 \\
                    & (0.84270, 0.97590) & (0.945100, 1.0000) \\
                \hline
                \multirow{2}{*}{p[Tom]}
                    & 0.75910 & 0.8575 \\
                    & (0.65060, 0.85800) & (0.745400, 0.9600) \\
                \hline
                \multirow{2}{*}{\(\psi\)}
                    & 0.87590 & 0.9126 \\
                    & (0.72910, 0.98880) & (0.746900, 0.9955) \\
                \hline
                \multirow{2}{*}{\(\theta\)}
                    & 0.24100 & 0.3779 \\
                    & (0.17890, 0.30900) & (0.286500, 0.4721) \\
                \hline
            \end{tabular}
        \end{center}
    \end{table}

    % Zumbro River!
    \begin{table}
        \caption{Zumbro River. The 402m data converged well. While
        \(\psi\) and \(\theta\) converged well for the 804m data, none of the
        other parameters did.}
        \label{Zumbro}
        \begin{center}
            \begin{tabular}{|l|l|l|}
                \hline
                \multicolumn{3}{|l|}{\textbf{Zumbro River, 2003}} \\
                \hline
                    & 402m & 804m \\
                \hline
                \multirow{2}{*}{E[Danny]}
                    & 0.02922 & 0.007066 \\
                    & (0.006786, 0.05994) & (0.0000, 0.03403) \\
                \hline
                \multirow{2}{*}{E[John]}
                    & 0.03388 & 0.008312 \\
                    & (0.008584, 0.06880) & (0.0000, 0.04188) \\
                \hline
                \multirow{2}{*}{E[Tom]}
                    & 0.02228 & 0.012960 \\
                    & (0.002430, 0.05385) & (0.0000, 0.05642) \\
                \hline
                \multirow{2}{*}{p[Danny]}
                    & 0.81890 & 0.837100 \\
                    & (0.749500, 0.88540) & (0.7615, 0.90700) \\
                \hline
                \multirow{2}{*}{p[John]}
                    & 0.81770 & 0.852500 \\
                    & (0.740100, 0.88670) & (0.7597, 0.92230) \\
                \hline
                \multirow{2}{*}{p[Tom]}
                    & 0.73120 & 0.805400 \\
                    & (0.625400, 0.82440) & (0.6885, 0.89780) \\
                \hline
                \multirow{2}{*}{\(\psi\)}
                    & 0.66550 & 0.651600 \\
                    & (0.549100, 0.78150) & (0.5098, 0.78210) \\
                \hline
                \multirow{2}{*}{\(\theta\)}
                    & 0.29950 & 0.395800 \\
                    & (0.236400, 0.36550) & (0.3100, 0.48520) \\
                \hline
            \end{tabular}
        \end{center}
    \end{table}

    Tables \ref{Mississippi}, \ref{Whitewater}, and \ref{Zumbro} contain our
    estimates for all of the data.

    With respect to the model's ability to converge on the actual data, the
    most important factors are plot size, the number of flights recorded, and
    the number of estimated parameters. In general, both plot size and the
    number of flights recorded must be sufficiently small and large,
    respectively, for the standard model with \(E\) to converge.

    The presence of numerous observations that occur in the second and third
    days since the most recent snow event allows one of the more crucial
    assumptions of our hierarchical model -- that otter tracks laid on each day
    after a snow event are visible throughout the entire observation period and
    therefore, that it is more likely for an observer to detect a track on
    later days -- to be thoroughly tested. Most of the \(\psi\) and \(\theta\)
    estimates of the data fall into ranges where our model is at its best, both
    with respect to convergence and to accurately estimating \(\psi\).

    As plot size increases, every estimated parameter tends to increase as
    well.  Although the vast majority of these increases, even at the 1608m
    plot size, lie well within the 95\% confidence interval of the accompanying
    estimate at the 402m plot size, this trend is intuitive. Once 402m plots
    are aggregated into 1608m plots, only one of the four 402m plots needs to
    be assigned a 1 in order for the entire 1608m plot to receive a 1. The
    amount of correlation between 1s determines the increase in the proportion
    of 1s to the number of plots -- as the amount of correlation increases,
    there will be more aggregated strings of 1s, so the increase in the
    proportion decreases.

    \subsection{Study design}

    Data with higher numbers of snow events and consecutive days since snow was
    not simulated because the model ran successfully on many combinations of
    one, two, and three observers, snow events, and consecutive days since
    snow.  Increasing the number of snow events or consecutive days since snow
    would also increase the total number of flights and therefore would not be
    more efficient than the combinations that were found to run successfully on
    the model.  Similarly, data with more than three observers was not
    simulated because it would also increase the number of flights, and there
    was success in estimating the parameters with only one, two, and three
    observers.

    The number of observers appears to be the least important of the three
    sampling variables.  Accurate estimates were obtained for all number of
    observers, as long as there were sufficient numbers of snow events and
    observations after each snow event.  For one observer, at least three snow
    events and three observations on consecutive days after each snow event are
    required for the model to converge and give accurate estimates.  For two or
    three observers, two events with three observations on consecutive days
    after each snow event is sufficient.

    The number of snow events appears to be the second-most important of the
    three sampling variables.  Although two snow events are required, only with
    one observer does the addition of a third snow event noticeably improve the
    model's estimates.

    The number of observations after each snow event appears to be the most
    important of the three sampling variables.  At least two observations after
    each snow event on consecutive days are required.  Additional observations
    after each event help the model converge more quickly, but the estimates
    become only slightly more accurate.

    Therefore, the most efficient combination of these sampling practices (in
    number of flights)  was found to be one observer, three snow events, and
    three consecutive days after each snow event.

\section{Discussion}

Our model performed surprisingly well, particularly given the complex and
sparse nature of the data analyzed. We were successfully able to incorporate
false detection, spatial data correlation, and the dependence of detection
probabilities on time since snowfall events into a model and still obtain both
accurate and precise results. While others have had success incorporating one
of these processes into statistical models \citep{Magoun2007, Royle2006,
Stanley2005}, we are not aware of any studies that have incorporated all three.
[Talk more about specifics when we get them.] Having the ability to model such
complex processes will make data collection in similar presence/absence surveys
both easier and less costly.

Interestingly, in our simulations with spatially correlated data, neither of
our two spatially explicit models performed particularly better than the
simpler spatially inexplicit model. While confidence intervals were sometimes
slightly smaller for the conditional autoregressive (CAR) models that took into
account the spatial correlation, the differences were small, and due to the
significantly longer computing times necessitated by the added complexity of
the spatial models, the simpler model may be a better choice. This is contrary
to what has been found in other studies that attempting to use CAR models with
similar binary data \citep{Wintle2006}.  However, the level of correlation in
their data may have been higher.

In addition, because funds for aerial surveys such as the ones used to collect
data in this study are limited, we explored different survey designs to
determine which aspects are most important in yielding precise and accurate
results. In particular, we examined numbers of observers, snow events, days
surveyed following snow events, and sites. [Make some sort of grand sweeping
conclusions about which are most important?] No single parameter appears most
important. However, we recommend a minimum sampling effort of [what?] to ensure
that the standard error for psi falls below [what?]. It is notable that if
there is more data available, the model will generally take less time to
converge.

These statistical methods could be used on data for other species, as well.
[John said we should talk about this. Which species?]

\section{Acknowledgements}

We would like to thank...

% Bibliography!
\bibliographystyle{Style}
\bibliography{Bibliography}

\section{Missing References}
\noindent

\textbf{[These are the ones that didn't show up in the paper.]}

\leftskip 0.15in

\parindent -0.15in

Gorman, T., Erb, J., McMillan, B.\ \& Martin, D.\ (2006) Space Use and
Sociality of River Otters ({\em Lontra Canadensis}) in Minnesota. {\em Journal
of Mammalogy}, {\bf 87}, 740--747.

\parindent -0.15in
Mackenzie, D.\ \& Royle, J.\ (2005) Designing Occupancy Studies: General Advice
and Allocating Survey Effort, {\em Journal of Applied Ecology}, {\bf 42},
1105-1114.

\parindent -0.15in
Sargeant, G., Sovada, M., Slivinski, C.\ \& Johnson, D.\ (2005) Markov Chain
Monte Carlo Estimation of Species Distributions: A Case Study of the Swift Fox
in Western Kansas, {\em Journal of Wildlife Management}, {\bf 69}, 483--497.

\end{document}